---
title: "project-draft"
author: "Alex Xiaotong Gui"
date: "3/1/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(skimr)
library(readr)
library(broom)
pokemon <- read_csv("pokemon.csv")
```

## Exploratory Analysis

```{r}
res<-cor(pokemon2)
library(corrplot)
corrplot(res, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```

```{r,fig.height=7,fig.width=4,warning=FALSE}
pokemon3<-pokemon %>%
  select(cp,cp_new,species,hp,weight,attack_strong_value,power_up_stardust)

library(GGally)
ggpairs(pokemon3,aes(col=species))
```


## Model Building

### preprocess the data for the model
```{r}
pokemon_model<-pokemon %>%
  select(-name,-notes) %>%
  select(-hp_new,-power_up_stardust_new,-attack_strong_value_new,-attack_weak_value_new,-weight_new,-height_new,-power_up_candy_new,-attack_strong_type_new,-attack_weak_type_new,-attack_strong_new,-attack_weak_new) %>%
  select(-attack_strong_type,-attack_weak_type,-attack_strong,-attack_weak)

#subset<-sample(c(TRUE,FALSE),nrow(pokemon_model),replace=TRUE,prob = c(1/3,2/3))
#poke.tst<-pokemon_model[subset,]
#poke.trn<-pokemon_model[!subset,]
#dim(poke.trn)
#dim(poke.tst)
#glimpse(poke.trn)
```

### best subset selection
```{r}
require(leaps)
poke.best <- regsubsets(cp_new~.,data=pokemon_model,nvmax = 10)
poke.best.sum<-summary(poke.best)
```

```{r}
best.cp <- which.min(poke.best.sum$cp) 
coef(poke.best, best.cp)
```

```{r}
best.bic <- which.min(poke.best.sum$bic)
coef(poke.best, best.bic)
```

```{r}
best.adjr <- which.max(poke.best.sum$adjr2)
coef(poke.best, best.adjr)
```

```{r}
bestsubset.lm1<-lm(cp_new~species+cp+hp+weight+power_up_stardust+power_up_candy+attack_strong_value,data=pokemon_model)
summary(bestsubset.lm1)
```

### Stepwise selection without interaction 
```{r}
step.lm.noint <- step(lm(cp_new~1, data = pokemon),cp_new~species+cp+hp+weight+power_up_stardust+power_up_candy+attack_strong_value+attack_weak_value, direction = "both")
```

```{r}
stepwise.lm.noint<-lm(cp_new ~ cp + species + attack_strong_value + hp + weight + power_up_stardust + 
    power_up_candy,data=pokemon_model)
summary(stepwise.lm.noint)
```

### stepwise selection with interaction
```{r}
step.lm.wint <- step(lm(cp_new~1, data = pokemon_model),cp_new~cp*species + attack_strong_value*species + hp*species + 
    weight*species + power_up_stardust*species + power_up_candy*species, direction = "both")
```


```{r}
stepwise.lm.wint<-lm(cp_new ~ cp + species + attack_strong_value + weight + cp:species + 
    species:attack_strong_value + species:weight,data=pokemon_model)
summary(stepwise.lm.wint)
```

### Use F test to test two nested model: with cp vs cp,hp,weight and attack strong interaction
```{r}
pokemon_full <- lm(cp_new~cp*species + attack_strong_value*species + hp*species + 
    weight*species + power_up_stardust*species + power_up_candy*species, data = pokemon_model)
anova(pokemon_full,stepwise.lm.wint)
```

```{r}
summary(step.lm3)
```


## Diagnostics

Identify outlying Y observations

```{r}
require(car)
outlierTest(stepwise.lm.wint)    #use a 0.05 significance level
pokemon_model[c(71,72,74),]
```

Outlying X observations
```{r}
nobs<-nrow(pokemon_model)
h.cutoff<-2*7/nobs
h.cutoff
hatvalues(stepwise.lm.wint)
plot(hatvalues(stepwise.lm.wint))
pokemon_model[hatvalues(stepwise.lm.wint)>h.cutoff,]
```

DFFITS
```{r}
dffit.cutoff<-2*sqrt(7/nobs)
plot(dffits(stepwise.lm.wint))
pokemon[abs(dffits(stepwise.lm.wint))>dffit.cutoff,]
```

DFBETAS
```{r}
dfbeta.cutoff<-2/sqrt(nobs)
dfbeta.cutoff
plot(dfbetas(stepwise.lm.wint))
pokemon[abs(dfbetas(stepwise.lm.wint))>dfbeta.cutoff,]
```


Cook's Distance
```{r}
pokemon_model[cooks.distance(stepwise.lm.wint)>=1,]
```

## Revist Model Building

###trying the data without the Eeves
```{r}
data_noEevees <- pokemon_model %>%
  filter(!species=="Eevee") %>%
  mutate(cp_new=log(cp_new),cp=log(cp))
```

Best Subset
```{r}
poke.best2 <- regsubsets(cp_new~.,data=data_noEevees,nvmax = 10)
poke.best.sum2<-summary(poke.best2)
poke.best.sum2
```
```{r}
best.cp2 <- which.min(poke.best.sum2$cp) 
coef(poke.best2, best.cp2)
summary(lm(cp_new~species+cp+hp+power_up_stardust+power_up_candy+attack_weak_value+attack_strong_value,data=data_noEevees))
```

```{r}
best.bic2 <- which.min(poke.best.sum2$bic)
coef(poke.best2, best.bic2)
bic.lm<-lm(cp_new~species+cp+power_up_stardust,data=data_noEevees)
summary(bic.lm)
```

```{r}
best.adjr2 <- which.max(poke.best.sum2$adjr2)
coef(poke.best2, best.adjr2)
```

```{r}
bestsubset.lm2<-lm(cp_new~species+cp+hp+weight+power_up_stardust+power_up_candy+attack_strong_value,data=data_noEevees)
summary(bestsubset.lm2)

bestsubset.lm3<-lm(cp_new~species+cp+power_up_stardust,data=data_noEevees)
summary(bestsubset.lm3)
```

```{r}
stepwise.lm2<- step(lm(cp_new~1, data = data_noEevees),cp_new~cp+species+hp+weight+power_up_stardust+power_up_candy+attack_strong_value+attack_weak_value, direction = "both")
summary(stepwise.lm2)
```

```{r}
anova(bic.lm,stepwise.lm2)
mean(vif(bic.lm))
```

### the entire dataset with log transformations

```{r}
pokemon_logged <- pokemon_model %>%
  mutate(cp_new=log(cp_new),cp=log(cp))
names(pokemon_logged)[2] <- c("cp_log")
names(pokemon_logged)[10] <- c("cp_new_log")
```

Best Subset
```{r}
poke.best.full <- regsubsets(cp_new_log~.,data=pokemon_logged,nvmax = 10)
poke.best.sum.full<-summary(poke.best.full)
poke.best.sum.full
```
```{r}
best.cp.full <- which.min(poke.best.sum.full$cp) 
coef(poke.best.full, best.cp.full)
summary(lm(cp_new_log~species+cp_log+hp+attack_strong_value,data=pokemon_logged))
```

```{r}
best.bic.full <- which.min(poke.best.sum.full$bic)
coef(poke.best.full, best.bic.full)
#bic.lm<-lm(cp_new~species+cp+power_up_stardust,data=data_noEevees)
#summary(bic.lm)
```

```{r}
best.adjr.full <- which.max(poke.best.sum.full$adjr2)
coef(poke.best.full, best.adjr.full)
summary(lm(cp_new_log~species+cp_log+power_up_stardust+power_up_candy+hp+weight+attack_strong_value,data=pokemon_logged))
```

##stepwise:
```{r}
stepwise.lm.final<- step(lm(cp_new_log~1, data = pokemon_logged),cp_new_log~cp_log+species+hp+weight+power_up_stardust+power_up_candy+attack_strong_value, direction = "both")
summary(stepwise.lm.final)
vif(stepwise.lm.final)
```

```{r}
stepwise.lm.full2<- step(lm(cp_new_log~1, data = pokemon_logged),cp_new_log~cp_log+cp_log*species+hp*species+weight*species+power_up_stardust*species+power_up_candy*species+attack_strong_value*species+attack_weak_value*species, direction = "both")
summary(stepwise.lm.full2)
vif(stepwise.lm.full2)
```

# check technical condition for the final model

# Y outlier
```{r}
require(car)
outlierTest(stepwise.lm.final)    #use a 0.05 significance level
pokemon_model[c(7,71,74),]
```

Outlying X observations
```{r}
nobs<-nrow(pokemon_logged)
h.cutoff<-2*7/nobs
h.cutoff
hatvalues(stepwise.lm.final)
plot(hatvalues(stepwise.lm.final))
pokemon_model[hatvalues(stepwise.lm.final)>h.cutoff,]
```

DFFITS
```{r}
dffit.cutoff<-2*sqrt(7/nobs)
plot(dffits(stepwise.lm.final))
pokemon[abs(dffits(stepwise.lm.final))>dffit.cutoff,]
```

DFBETAS
```{r}
dfbeta.cutoff<-2/sqrt(nobs)
dfbeta.cutoff
plot(dfbetas(stepwise.lm.final))
pokemon[abs(dfbetas(stepwise.lm.final))>dfbeta.cutoff,]
```

Cook's Distance
```{r}
pokemon_model[cooks.distance(stepwise.lm.final)>=1,]
```

```{r}
summary(stepwise.lm.final)
mean(vif(stepwise.lm.final))
```

# Residual plot final
```{r}
res<-resid(stepwise.lm.final)
plot(y=res,x=pokemon_logged$cp_new_log,ylab="Residuals",xlab="cp_new_log")
abline(0,0)
```

# Nested F test
```{r}
model1<-stepwise.lm.full
model2<-lm(cp_new ~ cp*species + species + attack_strong_value*species + hp*species + 
    weight*species,data=data_trans)
anova(model1,model2)
```

# Partial Determination

1) cp conditioned on other
```{r}
cond_cp<-anova(lm(cp_new~species+attack_strong_value+hp+weight+cp,data=data_trans))
cond_cp
cond_cp[5,2]/(cond_cp[5,2]+cond_cp[6,2])
```

2) species
```{r}
cond_species<-anova(lm(cp_new~cp+attack_strong_value+hp+weight+species,data=data_trans))
cond_species
cond_species[5,2]/(cond_species[5,2]+cond_species[6,2])
```

3) attack_strong
```{r}
cond_att<-anova(lm(cp_new~cp+species+hp+weight+attack_strong_value,data=data_trans))
cond_att
cond_att[5,2]/(cond_att[5,2]+cond_att[6,2])
```

4) hp
```{r}
cond_hp<-anova(lm(cp_new~cp+species+weight+attack_strong_value+hp,data=data_trans))
cond_hp
cond_hp[5,2]/(cond_hp[5,2]+cond_hp[6,2])
```

5) weight
```{r}
cond_weight<-anova(lm(cp_new~cp+species+hp+attack_strong_value+weight,data=data_trans))
cond_weight
cond_weight[5,2]/(cond_weight[5,2]+cond_weight[6,2])
```

## Confidence and Prediction Intervals

CI of coefficients
```{r}
require(broom)
tidy(stepwise.lm.full,conf.int = TRUE,conf.level = 0.95)
```



CI and PI of response:
```{r,warning=FALSE}
library(tidyr)
newdata=data.frame(species=c("Pidgey"),cp=c(log(200)),attack_strong_value=c(25),hp=c(41),weight=c(3))
pokemon_pred<-augment(stepwise.lm.full,newdata=newdata,type.predict = "response")
se.pred<-sqrt(glance(stepwise.lm.full)$sigma^2+pokemon_pred$.se.fit)
crit_val<-qt(.975,glance(stepwise.lm.full)$df.resid)
pred<-pokemon_pred %>%
  mutate(lower_CI=.fitted-crit_val*.se.fit,upperCI=.fitted+crit_val*.se.fit,lower_PI=.fitted-crit_val*se.pred,upper_PI=.fitted+crit_val*se.pred)
pred
```




